\textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\ {\footnotesize Bachelorarbeit}

### 2023-05-08 Paper/chapters/implementation-strategies.tex
camellia galois affine isomorphism

Bitslicing refers to the technique of splitting up $n$ bits into $m$ slices to
achieve a more efficient representation to operate on. The structure of
\texttt{GIFT} naturally offers possibilities for bitslicing. We split the
cipher state bits $b_{63}b_{62}\dots b_0$ into four slices $S_i,
i\in\{0,1,2,3\}$ such that the $i$-th slice contains all $i$-th bits of the
individual S-boxes. This is equivalent to transposing the bit matrix.

@@

This representation offers multiple advantages. We first note that computation
of the S-box can be executed in parallel, similar to the \texttt{vperm}
technique above. This can be done by finding an algorithmic way to apply the
S-box which has already been proposed by the original \texttt{GIFT} authors:

@@

\begin{description}
    \item[XOR cancellation property in key schedule:]
        \begin{align*}
            K_A&\leftarrow K_L\oplus K_R \\
            K_A&\leftarrow FE(FE(K_A, \Sigma_0=\mathrm{0xa09e667f3bcc908b}), \Sigma_1=\mathrm{0xb67ae8584caa73b2}) \\
            K_A&\leftarrow K_A\oplus K_L \\
        \end{align*}

@@

    \item[Efficiently computing $P\circ S$:] This technique applies to 64-bit processors.
        By preparing tables

@@

        , we can compute $(P\circ S)(X_{(64)})$ with only 8 table lookups and 7
        XORs in the following way:

@@

>>>>

Bitsliced implementation techniques were first introduced to improve the performance
of DES in 1997 and work by viewing a processor with $n$-bit registers as a
machine capable of executing $n$ bitwise operations at
once\cite{desslicing:1997}. Bitslicing offers a performance advantage by
splitting up $n$ bits into $m$ slices to achieve a more efficient
representation which can exploit this bitwise parallelism. The structure of
\texttt{GIFT} naturally offers possibilities for bitslicing. We split the
cipher state bits $b_{63}b_{62}\dots b_0$ into four slices $S_i,
i\in\{0,1,2,3\}$ such that the $i$-th slice contains all $i$-th bits of the
individual S-boxes. This is equivalent to transposing the bit matrix.

@@

This representation offers multiple advantages. We first note that computation
of the S-box can be executed in parallel, similar to the \texttt{vperm}
technique above. This can be done by finding a bitwise instruction sequence to
apply the S-box which has already been proposed by the original \texttt{GIFT}
authors:

@@

\begin{description}
    \item[XOR cancellation property in key schedule:]
        While deriving $K_A$ in the key schedule, the instruction sequence

        \begin{align*}
            K_A&\leftarrow K_L\oplus K_R \\
            K_A&\leftarrow FE(FE(K_A, \Sigma_0), \Sigma_1) \\
            K_A&\leftarrow K_A\oplus K_L \\
        \end{align*}

@@

    \item[Efficiently computing $(P\circ S)$:] This technique applies to 64-bit processors.
        By preparing tables

@@

        , we can compute $(P\circ S)(X_{(64)})$ using only 8 table lookups and 7
        XORs in the following way:

@@

\subsection{Byteslicing}

A byte-sliced implementation strategy for ARMv8 can be derived from already
existing x86-optimized implementations utilizing the AES-NI advanced encryption
standard instruction set \cite{bcfastimplx86:2013}. NEON itself possesses
cryptographic extensions for finite field arithmetic and AES as well as SHA
calculations. These can be used to produce an accelerated Camellia implementation
due to the algebraic similarity of the AES- and Camellia S-boxes.

\subsubsection{Hardware-accelerated Camellia S-box}
